On centrality of orthogonal $\Kt$

### 2016-05-08 paper.tex
Rewritten the introduction

\section*{Introduction}
The main goal of the present paper is to establish Quillen's local-global principle and the centrality of orthogonal $\Kt$ for arbitrary commutative ring $R$.

Recall that to each reduced root system $\Phi$ and a commutative ring $R$ one can associate a split simple group $\GG(\Phi, R)$ called a \emph{Chevalley group}.
For $\Phi$ of rank $\geq 2$ the abstract group $\GG(\Phi, R)$ contains a large normal subgroup $\E(\Phi, R)$ generated by the elementary root unipotents $t_\alpha(\xi)$, $\xi\in R$, $\alpha\in \Phi$, see~\cite{Ta}.
The quotient group $\Ko(\Phi, R)=\GG(\Phi, R)/\E(\Phi, R)$ is called the \emph{unstable $\Ko$-group}
This group shares many properties with its stable counterpart $\mathrm{SK}_1(R) = \lim\limits_{l\to\infty}\Ko(\rA_\ell, R)$, which, in turn,
is an essential direct summand of the $\Ko$-group ($\Ko(R) = \mathrm{SK}_1(R) \oplus R^*$).

Similarly to the definition of the algebraic $\Kt$-functor, one can define $\Kt(\Phi, R)$ as the kernel of the canonical projection $\phi\colon\St(\Phi, R)\twoheadrightarrow \E(\Phi, R)$.
Here $\St(\Phi, R)$ stands for the \emph{Steinberg group} of type $\Phi$, i.\,e. the group defined by formal generators $x_\alpha(\xi)$ modeling root unipotents $t_\alpha(\xi)$
and the explicit list of relations called Steinberg relations (see~\cref{sec:prelim}).

It is classically known that $\Kt(\Phi, R)$ is a central subgroup of $\St(\Phi, R)$ if $R$ is a commutative local ring and $\Phi$ has rank $\geq 2$, see \cite{St73}.
We refer to the latter statement as the ``centrality of $\Kt$''.
In~\cite{vdK} using the technique called ``another presentation'' W. van~der~Kallen has shown that the centrality of $\Kt$ holds for $\Phi=\rA_\ell$, $\ell\geq 3$ and arbitrary commutative ring $R$.
Using similar ideas Tulenbaev in~\cite{Tul0} further generalised van~der~Kallen's result to associative rings that are module-finite over their centres.

Thus, it is natural to expect that the centrality of $\Kt$ holds for arbitrary commutative rings and different types of Chevalley groups, as well.
In \cite{Lav}, \cite{SCh} the authors of the present paper have established the centrality of $\Kt(\rC_\ell, R)$ for $\ell\geq 3$ and $\Kt(\rE_\ell, R)$, $\ell=6,7,8$.
In turn, the key result of this paper is the following theorem.
\begin{tm}\label{tm:centrality}  Let $R$ be an arbitrary commutative ring and $\ell\geq 4$ then 
the group $\Kt(\rD_\ell, R)$ is contained in the centre of $\St(\rD_\ell, R)$. \end{tm}

\cref{tm:centrality} is essentially a corollary of the following local-global principle.
\begin{tm}\label{tm:lg-principle} Let $\Phi$ be an irreducible simply-laced root system of rank $\geq 3$ and $R$ be arbitrary commutative ring.
Then an element $g\in\St(\Phi,\,R[X])$ satisfying $g(0)=1\in\St(\Phi,\,R)$ is trivial in $\St(\Phi,\,R[X])$
if and only if the images $g_M\in\St(\Phi,\,R_M[X])$ of $g$ under localisation morphisms $R\to R_M$ are trivial for all maximal ideals $M\trianglelefteq R$. \end{tm}

@@

The Steinberg group $\St(n, R) = \St(\rA_{n-1}, R)$ is the group presented by generators
$x_{ij}(r)$, $1\leq i\neq j\leq n$, $r\in R\}$ and relations
\setcounter{equation}{0}
\renewcommand{\theequation}{S\arabic{equation}}
\begin{align}
&x_{ij}(r)x_{ij}(s)=x_{ij}(r+s), \label{add0}\\
&[x_{ij}(r),\,x_{hk}(s)]=1,\text{ for }h\neq j,\ k\neq i, \label{ccf1}\\
&[x_{ij}(r),\,x_{jk}(s)]=x_{ik}(rs) \label{ccf2}.
\end{align}
The natural projection $\phi\colon\St(n,\,R)\rightarrow\E(n,\,R)$ sends $x_{ij}(r)$ to $t_{ij}(r)$.

@@

\section{Patching Tulenbaev maps (to be written)}\label{sec:patching}

The next section is devoted to the proof of Theorems~\ref{tm:centrality}--\ref{tm:lg-principle} for Steinberg groups corresponding to 
arbitrary simply-laced root system of rank $l\geq3$.
To do so, we glue maps $T$ constructed in the previous section following the approach of~\cite{SCh}.

\printbibliography

>>>>

\section*{Introduction}
The aim of this paper is to establish Quillen's local-global principle and the centrality of even orthogonal $\Kt$ for arbitrary commutative ring $R$.

Recall that to each reduced irreducible root system $\Phi$ and a commutative ring $R$ one can associate a split simple group $\GG(\Phi, R)$ called a \emph{Chevalley group}.
The {\it elementary group} $\E(\Phi, R)$ is defined as the subgroup of $\GG(\Phi, R)$ generated by elementary root unipotents $t_\alpha(\xi)$, $\xi\in R$, $\alpha\in \Phi$.
By definition, the Steinberg group $\St(\Phi, R)$ is presented by generators $x_\alpha(\xi)$, which model the unipotents $t_\alpha(\xi)$, subjected to Steinberg relations (see~\cref{sec:prelim}). 
There is a natural map $\phi\colon\St(\Phi, R)\twoheadrightarrow \GG(\Phi, R)$ sending $x_\alpha(\xi)$ to $t_\alpha(\xi)$ whose image equals $\E(\Phi, R)$.

By a theorem of Taddei (see~\cite{Ta}) the elementary group $\E(\Phi, R)$ is a normal subgroup of $\GG(\Phi, R)$ if $\Phi$ has rank at least $2$.
Thus, similarly to the definition of algebraic $\mathrm{SK}_1$ and $\Kt$-functors, one can define the groups $\Ko(\Phi, R)$, $\Kt(\Phi, R)$ as the cokernel and kernel of $\phi$:
$$\xymatrix{1 \ar[r] & \Kt(\Phi, R) \ar[r] & \St(\Phi, R) \ar[r] & \GG(\Phi, R) \ar[r] & \Ko(\Phi, R) \ar[r] & 1.}$$

It is a classical theorem of M.~Kervaire and R.~Steinberg that the stable Steinberg group $\St(R)$ is the universal central extension of $\E(R)$ for arbitrary $R$, see \cite[Theorem~III.5.5]{Kbook}.
The unstable analogue of the ``universality'' part of this theorem was obtained by M.~Stein and W.~van~der~Kallen in 1970's, see \cite{St1, KaS}.
However, until recently, the analogue of the ``centrality'' part of the theorem was only known for $\Phi=\rA_\ell$, $\ell\geq 4$, see~\cite[Corollary~2]{vdK} and for other classical $\Phi$ ``in the stable range'',
i.\,e. when the dimension of $R$ is suffiently small with respect to the rank of $\Phi$, see~\cite[Corollary~3.4]{St78}.

Recently the authors of this paper have established the centrality of $\Kt(\rC_\ell, R)$ ($\ell\geq 3$) and $\Kt(\rE_\ell, R)$ ($\ell=6,7,8$) for arbitrary commutative ring $R$, see \cite{Lav, SCh}.
In turn, the key result of this paper is the following theorem.
\begin{tm}\label{tm:centrality}  Let $R$ be an arbitrary commutative ring and $\ell\geq 4$ then 
the group $\Kt(\rD_\ell, R)$ is a central subgroup of $\St(\rD_\ell, R)$. \end{tm}

\cref{tm:centrality} follows from the following local-global principle.
We refer the reader to the proof ~\cite[Theorem~]{SCh}
\begin{tm}\label{tm:lg-principle} Let $\Phi$ be an irreducible simply-laced root system of rank $\geq 3$ and $R$ be arbitrary commutative ring.
Then an element $g\in\St(\Phi,\,R[X])$ satisfying $g(0)=1\in\St(\Phi,\,R)$ is trivial in $\St(\Phi,\,R[X])$
if and only if the images $g_M\in\St(\Phi,\,R_M[X])$ of $g$ under localisation morphisms $R\to R_M$ are trivial for all maximal ideals $M\trianglelefteq R$. \end{tm}

@@

The Steinberg group $\St(n, R) = \St(\rA_{n-1}, R)$ is the group defined by generators
$x_{ij}(r)$, $1\leq i\neq j\leq n$, $r\in R\}$ and relations
\setcounter{equation}{0}
\renewcommand{\theequation}{S\arabic{equation}}
\begin{align}
&x_{ij}(r)x_{ij}(s)=x_{ij}(r+s), \label{add0}\\
&[x_{ij}(r),\,x_{hk}(s)]=1,\text{ for }h\neq j,\ k\neq i, \label{ccf1}\\
&[x_{ij}(r),\,x_{jk}(s)]=x_{ik}(rs) \label{ccf2}.
\end{align}
The natural projection $\phi\colon\St(n,\,R)\rightarrow\E(n,\,R)$ sends $x_{ij}(r)$ to $t_{ij}(r)$.

@@

%\printbibliography
